Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine

### 2016-01-14 paper.tex
Wosnik edits part 2

In conjunction with the cross-flow turbine blade kinematics, the foil
coefficients were used to approximate turbine performance by calculating the
peak torque coefficient on the upstream half of the blade path. The turbine
torque coefficient $C_T$ can be related to the blade section chordwise force
coefficient $C_c$ by
\begin{equation}
C_T = \frac{C_c c}{2R} \frac{|U_\mathrm{rel}|^2}{U_\infty^2},
\label{eq:ct}
\end{equation}
where the blade section chordwise force (for zero preset pitch) is given by
\begin{equation}
C_c = C_l \sin \alpha - C_d \cos \alpha.
\label{eq:cc}
\end{equation}

@@

From the peak torque coefficient metric plotted in
Figure~\ref{fig:foils-C_T-Re-dep}, the Reynolds number independence is achieved
at lower values, and more dramatically. We see that the trend of the
(non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1
\times 10^5$, but then continues to increase linearly with a small slope with
$Re$. This is not matched by the experimental data, which at higher $Re$ looks
more like the cambered foil results. Though this method is very approximate, it
nevertheless appears to be a reasonable way to predict the transitional Reynolds
number regime for cross-flow turbine performance using only 2-D static airfoil
characteristics

@@

Spectra of the cross-stream velocity normalized by the free stream were
computing using a fast Fourier transform-based method, applying a Hanning window
and averaging over four adjacent frequency bands to decrease confidence
intervals. These spectra are plotted in Figure~\ref{fig:wake-spectra} for
regions on either side of the turbine. On the $-y$ side of the turbine there is
broadband turbulence produced by blade stall, and on the $+y$ side there is a
clear peak in the spectra caused by the blade passage. We see that on both sides
there is higher spectral energy at lower Reynolds numbers. On the $+y$ side of
the turbine we notice higher spectral energy at the blade passage frequency's
first harmonic, or $6 f_\mathrm{turbine}$. This is likely due to the blade's
shed vorticity being less stable at higher $Re$.

@@

Results from the terms that can be computed from the experimental data (i.e.
excluding $\partial / \partial x$) from the right hand sides of
Equations~\ref{eq:momentum} and \ref{eq:K-full} are plotted in
Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph}, respectively.
Derivatives were computed using a second order finite difference scheme, with
central differencing for interior points, and inward facing schemes for the
boundaries. A weighted average for each term was then calculated based on the
grid spacing. We note that similar to \cite{Bachant2015-JoT}, the vertical
advection at this point in the wake is the dominant contributor to positive wake
recovery, caused by the unique vortex pattern created by the blade tips and
wakes.

We see that in general, levels of turbulent transport are slightly lower at
larger Reynolds numbers. The viscous diffusion and dissipation, though still
three orders of magnitude smaller than the other terms, do increase at low
Reynolds numbers, which is consistent with the physical meaning of the Reynolds
number itself. From these results one might expect that the viscous effects will
become significant to the wake dynamics around $Re_D \sim 10^4$.

Transport due to cross-stream advection appears to become more negative at
higher $Re$. This could be a consequence of increasing free surface deformation
(higher $Fr$), which then decreases the effective flow cross-sectional area
downstream of the turbine, forcing more flow to accelerate around the sides of
the turbine. Note that we also see slightly higher drag coefficients at higher
$Re$ (potentially for the same reason), which also helps explain larger negative
values of cross-stream advection.

@@

In this study it was demonstrated that the performance of a high solidity
($c/R=0.28$) cross-flow turbine becomes essentially $Re$-independent at a
Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
10^5$. The power coefficient values on the left side of the $C_P$--$\lambda$
curve are further reduced at low $Re$ due to deeper stalling. The
$Re$-independent threshold corresponds to that at which the foil suction surface
boundary layer becomes turbulent before having to recover static pressure
against the adverse pressure gradient, highlighted in \cite{Lissaman1983,
    McMasters1980, Carmichael1981}.

@@

In the near-wake of the turbine we observed lower levels of turbulence with
increasing Reynolds number, along with lower levels of turbulent transport with
respect to mean streamwise momentum and mean kinetic energy recovery in the
streamwise direction. Vertical advection---the largest transport mechanism
measured---showed little $Re$-dependence, whereas the negative effects of
cross-stream advection were enhanced at high $Re$, which may be due to larger
free surface deformation reducing the near-wake's cross-sectional area. Despite
being orders of magnitude smaller than other transport processes, viscous
diffusion increased rapidly with decreasing $Re$, and is expected to become
significant to the wake dynamics around $Re_D \sim 10^4$. Overall, the total
wake transport measured leveled off at essentially the same Reynolds number that
the performance did---$Re_D \approx 0.8 \time 10^6$.

From these results we recommend that physical model tests be performed at $Re_D
\sim 10^6$ to provide reasonable predictions of full-scale performance. This
threshold also applies to the validation of predictive engineering models,
especially for high-fidelity CFD models where the boundary layer is to be
resolved, since transition to turbulence plays an important role in overall
blade loading. Note that in our case blockage, though reasonably low, may be
increasing flow through the turbine compared with a free case, which could mean
$Re$ thresholds for free flows could be slightly higher, though a reliable
blockage correction algorithm has not yet been agreed upon for CFTs
\cite{Cavagnaro2014}.

>>>>

In conjunction with the cross-flow turbine blade kinematics, the foil
coefficients were used to approximate turbine performance by calculating the
peak torque coefficient on the upstream half of the blade path. The turbine
torque coefficient $C_T$ can be related to the blade section chordwise force
coefficient $C_c$ by
\begin{equation}
C_T = \frac{C_c c}{2R} \frac{|U_\mathrm{rel}|^2}{U_\infty^2},
\label{eq:ct}
\end{equation}
where the blade section chordwise force coefficient (for zero preset pitch) is
given by
\begin{equation}
C_c = C_l \sin \alpha - C_d \cos \alpha.
\label{eq:cc}
\end{equation}

@@

From the peak torque coefficient metric plotted in
Figure~\ref{fig:foils-C_T-Re-dep}, the Reynolds number independence is achieved
at lower values, and more dramatically. We see that the trend of the
(non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1
\times 10^5$, but then continues to increase slowly and linearly with $Re$. This
is not matched by the experimental data, which at higher $Re$ looks more like
the cambered foil results. Though this method does not provide absolute
predictions of turbine performance, it predicts the transitional Reynolds number
regime for cross-flow turbine performance using only 2-D static airfoil
characteristics, i.e., Reynolds number scaling of the peak $C_T$ computed this
way behaves much like that of the measured turbine performance.

@@

Spectra of the cross-stream velocity normalized by the free stream were
computed using a fast Fourier transform-based method, applying a Hanning window
and averaging over four adjacent frequency bands to decrease confidence
intervals. These spectra are plotted in Figure~\ref{fig:wake-spectra} for
regions on either side of the turbine. On the $-y$ side of the turbine there is
broadband turbulence produced by blade stall, and on the $+y$ side there is a
clear peak in the spectra caused by the blade passage. We see that on both sides
there is higher spectral energy at lower Reynolds numbers. On the $+y$ side of
the turbine we notice higher spectral energy at the blade passage frequency's
first harmonic, or $6 f_\mathrm{turbine}$. This is likely due to the blade's
shed vorticity being less stable at higher $Re$.

@@

Results from the terms that can be computed from the experimental data (i.e.
excluding $\partial / \partial x$) from the right hand sides of
Equations~\ref{eq:momentum} and \ref{eq:K-full} are plotted in
Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph}, respectively.
Derivatives were computed using a second order finite difference scheme, with
central differencing for interior points, and inward facing schemes for the
boundaries. A weighted average for each term was then calculated based on the
grid spacing. We note that similar to \cite{Bachant2015-JoT}, the vertical
advection at this point in the wake is the dominant contributor to positive wake
recovery, caused by the unique vortex pattern created by the blade tips and
blade wakes.

We see that in general, levels of turbulent transport are slightly lower at
larger Reynolds numbers. The viscous diffusion and dissipation, though still
three orders of magnitude smaller than the other terms, do increase at low
Reynolds numbers, which is consistent with the physical meaning of the Reynolds
number itself. From these results one might expect that the viscous effects will
become significant to the wake dynamics as Reynolds number is decreased below
$Re_D \sim 10^4$.

Transport due to cross-stream advection appears to become more negative at
higher $Re$. This could be a consequence of increasing free surface deformation
(higher $Fr$), which then decreases the effective flow cross-sectional area
downstream of the turbine, forcing more flow to accelerate around the sides of
the turbine. Note that we also observe slightly higher drag coefficients at
higher $Re$ (potentially for the same reason), which also helps explain larger
negative values of cross-stream advection.

@@

In this study it was demonstrated that the performance of a high solidity
($c/R=0.28$) cross-flow turbine becomes essentially $Re$-independent at a
Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
10^5$. The power coefficient values on the left hand side of the
$C_P$--$\lambda$ curve are reduced more than those on the right hand side at low
$Re$ due to deeper stalling. The $Re$-independent threshold corresponds to that
at which the foil suction surface boundary layer becomes turbulent before having
to recover static pressure against the adverse pressure gradient, highlighted in
\cite{Lissaman1983, McMasters1980, Carmichael1981}.

@@

In the near-wake of the turbine we observed lower levels of turbulence with
increasing Reynolds number, along with lower levels of turbulent transport with
respect to mean streamwise momentum and mean kinetic energy recovery in the
streamwise direction. Vertical advection---the largest transport mechanism
measured---showed little $Re$-dependence, whereas the negative effects of
cross-stream advection were enhanced at high $Re$, which may be due to larger
free surface deformation effectively reducing the near-wake's cross-sectional
area. Despite being orders of magnitude smaller than other transport processes,
viscous diffusion increased rapidly with decreasing $Re$, and is expected to
become significant to the wake dynamics around $Re_D \sim 10^4$. Overall, the
total wake transport measured leveled off at essentially the same Reynolds
number that the performance did---$Re_D \approx 0.8 \times 10^6$.

From these results we recommend that physical model tests of cross-flow turbines
be performed at $Re_D \sim 10^6$ or greater to provide reasonable predictions of
full-scale performance. This threshold also applies to the validation of
predictive engineering models, especially for high-fidelity CFD models where the
boundary layer is to be resolved, since transition to turbulence plays an
important role in overall blade loading. Note that in our case blockage, though
reasonably low, may be increasing flow through the turbine compared with a free
case, which could mean $Re$ thresholds for free flows could be slightly higher,
though a reliable blockage correction algorithm has not yet been agreed upon for
CFTs \cite{Cavagnaro2014}.
